---
editor_options: 
  chunk_output_type: console
---

## Add data Layers

Spatial objects (points, lines, polygons, rasters) in your R environment can also be added as map layers, provided that they have a CRS defined with a datum. Leaflet will try to make the necessary transformation to display your data in [EPSG:3857](https://epsg.io/3857). 


===

Read in data using [sf](){:.rlib} and [raster](){:.rlib} packages. Points are from the 
[Water Quality Portal](https://agupubs.onlinelibrary.wiley.com/doi/full/10.1002/2016WR019993) 
accessed via the [dataRetrieval](){:.rlib} package using [this script]((https://github.com/SESYNC-ci/leaflet-in-R-lesson/blob/master/get_wqp_sites.R)).
County boundaries are from the [Census](https://www.census.gov/geo/maps-data/data/tiger-line.html) 
and [Watershed boundaries](https://water.usgs.gov/maps.html) are from USGS.

```{r read_data, handout = 1, message = FALSE, results='hide'}
library(sf)
library(raster)
library(dplyr)
wqp_sites <- st_read("data/wqp_sites")
counties_md <- st_read("data/cb_2016_us_county_5m") %>% 
  filter(STATEFP == "24") %>% st_transform(4326)
wbd_reg2 <- st_read("data/huc250k/") %>%
  filter(REG == "02") %>% st_transform(4326)
nlcd <- raster("data/nlcd_crop.grd")
```

Here we use `filter` to subset the county boundaries to the state with code `24` (Maryland) 
and the watershed boundaries to hydrologic unit `02` (Mid-Atlantic). We transform both
to [EPSG:4326](https://epsg.io/4326) (latitude-longitude coordinates with WGS84 datum).
{:.notes}

===

Add a polygons layer, being sure to specify the `data = ` argument. All of the `add*` layers have "map" as the first argument, facilitating use of the pipe (`%>%`) operator, but `data` is not the second argument so it must be named. 

```{r leaflet_polygons, handout = 1}
leaflet() %>%
  addTiles() %>%
  addPolygons(data = counties_md)
```

===

Customize Polygons with border and fill colors and opacities.

```{r leaflet_polygonscolor, handout = 1}
leaflet() %>%
  addTiles() %>%
  addPolygons(data = counties_md, 
              color = "green", 
              fillColor = "gray", 
              fillOpacity = 0.75, 
              weight = 1)
```

===

Attributes can be referred to using formula syntax `~` to access attribute data from an `sf` object or `Spatial*DataFrame`. What is the difference between a popup and a label?

```{r leaflet_popvhov, handout = 1}
leaflet() %>%
  addTiles() %>%
  addPolygons(data = counties_md, popup = ~NAME, group = "click") %>%
  addPolygons(data = counties_md, label = ~NAME, group = "hover") %>%
  addLayersControl(baseGroups = c("click", "hover"))
```

Notice that the names `"click"` and `"hover"` that we assigned to the groups correspond to the different behavior of popups and labels!
{:.notes}

===

For numeric or categorical data, create a color palette from [colorbrewer](http://colorbrewer2.org/#) using `colorNumeric()`, `colorFactor()`, `colorBin()`, or `colorQuantile()`. 

```{r leaflet_pals, handout = 1}
pal <- colorNumeric("PiYG", wbd_reg2$AREA)
leaflet() %>%
  addTiles() %>%
  addPolygons(data = wbd_reg2, 
              fillColor = ~pal(AREA), 
              fillOpacity = 1, 
              weight =1)
```

Load the RColorBrewer library to see information about palettes in the `brewer.pal.info` data.frame.

Leaflet's color palette syntax can be a little confusing. The functions called `color*()` return another function, so here the object `pal` is a custom-defined function that maps a numeric value to a color. In this case the range of values in `wbd_reg2$AREA` is mapped to the range of colors in the `"PiYG"` palette. Then we use formula notation, `~pal(AREA)`, to tell `addPolygons()` that the `fillColor` should be computed for each polygon by applying the function `pal()` to the `AREA` value of that polygon.
{:.notes}

===

Add a legend (which will likely duplicate arguments from the layer that used that palette). 

```{r leaflet_legend, handout = 1}
leaflet() %>%
  addTiles() %>%
  addPolygons(data = wbd_reg2, 
              fillColor = ~pal(AREA), 
              fillOpacity = 0.9, 
              weight =1) %>%
  addLegend(pal = pal, values = wbd_reg2$AREA, title = "Watershed Area")
```

===

Add points as markers, circle markers, or [customized icons](https://rstudio.github.io/leaflet/markers.html).

```{r leaflet_pts, handout = 1}
leaflet() %>%
  addTiles() %>%
  addMarkers(data = wqp_sites[1:1000,], group = "markers") %>%
  addCircleMarkers(data = wqp_sites[1:1000,], radius = 5, group = "circlemarkers") %>%
  addLayersControl(baseGroups = c("markers", "circlemarkers"))
```

Here the `radius` argument indicates the size of the `"circlemarkers"` group.
{:.notes}

===

Display a different number of markers at each zoom level using `clusterOptions` on markers or circle markers layers.

```{r leaflet_cluster, handout = 1}
leaflet() %>%
  addTiles() %>%
  addMarkers(data = wqp_sites, clusterOptions = markerClusterOptions(),
             popup = ~paste(MntrnLN, ",", OrgnztI))
```

Here, we use the function `markerClusterOptions()` with all default arguments. This will use pre-defined logic to decide 
how many markers to collapse together at each zoom level. The `popup` argument will only
show the popup text when you are zoomed in far enough to click on an individual marker.
{:.notes}

===

Rasters can be slow to load. Be patient but also check out the `maxBytes =` argument if need be. 

```{r leaflet_raster, handout = 1}
leaflet() %>%
  addTiles() %>%
  addRasterImage(nlcd)
```