---
---

## Exercises

===

### Exercise 1

Create a leaflet map with a marker at a location and zoom level of your choice that includes

- Background imagery from a tile provider of your choice (**Hint**: browse options in the `providers` list)
- A label that appears when the mouse hovers over the marker (**Hint**: use the `label` argument to `addMarkers()`)

[View solution](#solution-1)
{:.notes}

===

### Exercise 2

Create a leaflet map with default background imagery and the `counties_md` polygons colored by area (the `ALAND` column), using the `"Oranges"` palette. 

**Hint**: You will need to define a palette function using `colorNumeric()` before you create the map. The first argument to `colorNumeric()` should be the name of the palette (`"Oranges"`), and the second argument is the range of values to map to colors (`counties_md$ALAND`).

[View solution](#solution-2)
{:.notes}

===

### Exercise 3

Modify the Shiny app on worksheet 3 (the app where the user selects a date of MODIS imagery to display) to do the following:

- Always display both the OpenStreetMap and MODIS layers, instead of letting the user choose.
- Include a slider labeled "Set Opacity" that will allow the user to modify the variable `input$opacity`.
- Use the variable `input$opacity` to set the opacity of the MODIS layer between 0 and 1, with the default value being 1.

This is challenging so don't be afraid to look at the hints!

**Hints**:

- Remove the `addLayersControl()` function entirely to disable the switching on and off of layers.
- Add another argument to `absolutePanel()` after the `dateInput()`. This argument will be `sliderInput()`.
- The arguments to `sliderInput()` should be the input variable name (`"opacity"`), the label that the user will see (`"Set Opacity"`), the minimum value (`min = 0`), the maximum value (`max = 1`), and the default value (`value = 1`).
- Finally, add another argument to `providerTileOptions()` to set the opacity (`opacity = input$opacity`).

[View solution](#solution-3)
{:.notes}

===

## Solutions

===

### Solution 1

This is an example solution centered on the Washington Monument, zoomed in to level 14, with `Stamen.Toner` background imagery.

```{r eval = FALSE}
lng <- -77.0351929
lat <- 38.8894791
leaflet() %>%
  setView(lng = lng, lat = lat, zoom = 14) %>%
  addProviderTiles(providers$Stamen.Toner) %>%
  addMarkers(lng = lng, lat = lat, label = "Yep, that's it") 
```

[Return](#exercise-1)
{:.notes}

===

## Solution 2

```{r eval = FALSE}
pal <- colorNumeric('Oranges', counties_md$ALAND)
leaflet() %>%
  addTiles() %>%
  addPolygons(data = counties_md, 
              fillColor = ~ pal(ALAND))
```

[Return](#exercise-2)
{:.notes}

===

### Solution 3

```{r eval = FALSE}
ui <- fluidPage(
  leafletOutput("map", height = 800),
  # add calendar panel
  absolutePanel(top = 100, right = 10, draggable = TRUE,
                dateInput("dateinput", "Imagery Date", value = "2018-03-28"),
                sliderInput("opacity", "Opacity", min = 0, max = 1, value = 1)
  ))

server <- function(input, output) {
  
  output$map <- renderLeaflet({
    # leaflet map is defined here with tiles and initial view, and no layer control
    leaflet() %>%
      addTiles(group = "OSM") %>%
      addProviderTiles(providers$NASAGIBS.ModisTerraTrueColorCR, group = "Modis",
                       options = providerTileOptions(time = input$dateinput, opacity = input$opacity)) %>%
      setView(lng = -76.505206, lat = 38.9767231, zoom = 7)
  })
  
}

shinyApp(ui, server)
```

[Return](#exercise-3)
{:.notes}
