---
---

## Publish a Leaflet Map

You can render a map and publish it using the `htmlwidgets` package.  
HTML widgets work just like R plots, except they produce interactive visualizations. 

===

A line or two of R code is all it takes to produce a Leaflet map as an HTML file.  
Your HTML map can be used as a standalone website or embedded in an existing website. 

===

Using leaflet, create a map centered on Annapolis and add a weather data tile. 
```{r leaflethtml1, handout = 1}
library(leaflet)
library(htmlwidgets)

map <- leaflet() %>%
  addTiles() %>%
  setView(lng = -76.505206, lat = 38.9767231, zoom = 5) %>%
  addWMSTiles(
    "http://mesonet.agron.iastate.edu/cgi-bin/wms/nexrad/n0r.cgi",
    layers = "nexrad-n0r-900913", 
    options = WMSTileOptions(format = "image/png", transparent = TRUE),
    attribution = "Weather data © 2012 IEM Nexrad"
  )
```

===

Now that we have a map, let's save it by using the `saveWidget()`function.
`saveWidget()` will save your map to your working directory as a `.html` file.
```{r leaflethtml2, handout = 1, message = FALSE, eval = FALSE}
saveWidget(map, file="map.html")
```

You can now open `map.html` in your browser as a full-screen html file.

===

---

To share your map, you can upload the `html` file to GitHub or any web server. 
We suggest you use GitHub Pages to share your map. 

To publish your map on GitHub Pages, do the following: 
1. Create a repository on GitHub.
2. Rename your `map.html` file to `index.html` and push this file to your repository. 
3. Go to your repository's settings and enable GitHub Pages. 
4. After a few minutes, your page will be ready to go and a link will be provided. 

Your map is now live. You can now share the link with your colleagues to share your map!

===

For more information about GitHub Pages, including a walkthrough of the steps above, 
please see our lesson on [Advanced git Techniques](https://cyberhelp.sesync.org/advanced-git-lesson/#/slides/pages). 
For information on getting started with GitHub, check out our [Basic git](https://cyberhelp.sesync.org/basic-git-lesson/) lesson.
{:.notes}
---

If you prefer to embed your map inside a project, personal, or any other website you can do so by using the html `<iframe>` tag.

```html
<iframe src="map.html" frameborder="0" width="50%" height="200px"></iframe>
```
This is a good option when you already have a site built and would like to add your map to share it.

===

Web development is out of the scope of this lesson, but please check this [tutorial](https://www.tutorialrepublic.com/html-tutorial/html-iframes.php) out if you would like to learn more about using the `iframe` tag in your websites. 
{:.notes}

===
