---
editor_options: 
  chunk_output_type: console
---

## Add data Layers

Spatial objects (points, lines, polygons, rasters) in your R environment can also be added as map layers, provided that they have a CRS defined with a datum. Leaflet will try to make the necessary trasnformation to display your data in [EPSG:3857](https://epsg.io/3857). 


===

Read in data using sf and raster packages. Points are from the [Water Quality Portal](https://agupubs.onlinelibrary.wiley.com/doi/full/10.1002/2016WR019993) [accessed via](https://github.com/SESYNC-ci/leaflet-in-R-lesson/blob/master/get_wqp_sites.R) the `dataRetrieval` package. County boundaries are from the [Census](https://www.census.gov/geo/maps-data/data/tiger-line.html) and [Watershed boundaries](https://water.usgs.gov/maps.html) are from USGS.

```{r read_data, title = "{{ site.handouts[1] }}"}
library(sf)
library(raster)
library(dplyr)
wqp_sites <- st_read("data/wqp_sites") 
counties_md <- st_read("data/cb_2016_us_county_5m") %>% 
  filter(STATEFP == "24") %>% st_transform(4326)
wbd_reg2 <- st_read("data/huc250k/") %>%
  filter(REG == "02") %>% st_transform(4326)
nlcd <- raster("data/nlcd_crop.grd")
```

===

Add a polygons layer, being sure to specify the `data = ` argument. All of the `add*` layers have "map" as the first argument, facilitating use of the pipe (`%>%`) operator, but `data` is not the second argument so it must be named. 

```{r leaflet_polygons, title = "{{ site.handouts[1] }}"}
leaflet() %>%
  addTiles() %>%
  addPolygons(data = counties_md)

```

===

Customize Polygons with border and fill colors and opacities.

```{r leaflet_polygonscolor, title = "{{ site.handouts[1] }}"}
leaflet() %>%
  addTiles() %>%
  addPolygons(data = counties_md, 
              color = "green", 
              fillColor = "gray", 
              fillOpacity = 0.75, 
              weight = 1)

```

===

Attributes can be referred to using formula syntax `~` to access attribute data from an `sf` object or `Spatial*DataFrame`. What is the difference between a popup and a label?

```{r leaflet_popvhov, title = "{{ site.handouts[1] }}"}
leaflet() %>%
  addTiles() %>%
  addPolygons(data = counties_md, popup = ~NAME, group = "click") %>%
  addPolygons(data = counties_md, label = ~NAME, group = "hover") %>%
  addLayersControl(baseGroups = c("click", "hover"))


```

===

For numeric or categorical data, create a color palette from [colorbrewer](http://colorbrewer2.org/#) using `colorNumeric()`, `colorFactor()`, `colorBin()`, or `colorQuantile()`. `brewer.pal.info` is a data.frame containing information about the palettes.

```{r leaflet_pals, title = "{{ site.handouts[1] }}"}
leaflet() %>%
  addTiles() %>%
  addPolygons(data = wbd_reg2, 
              fillColor = ~pal(AREA), 
              fillOpacity = 1, 
              weight =1)

```

===

Add a legend (which will likely duplicate arguments from the layer that used that palette). 

```{r leaflet_legend, title = "{{ site.handouts[1] }}"}
leaflet() %>%
  addTiles() %>%
  addPolygons(data = wbd_reg2, 
              fillColor = ~pal(AREA), 
              fillOpacity = 0.9, 
              weight =1) %>%
  addLegend(pal = pal, values = wbd_reg2$AREA, title = "Watershed Area")

```

===

Add points as markers, circle markers, or [customized icons](https://rstudio.github.io/leaflet/markers.html).

```{r leaflet_pts, title = "{{ site.handouts[1] }}"}
leaflet() %>%
  addTiles() %>%
  addMarkers(data = wqp_sites[1:1000,], group = "markers") %>%
  addCircleMarkers(data = wqp_sites[1:1000,], radius = 5, group = "circlemarkers") %>%
  addLayersControl(baseGroups = c("markers", "circlemarkers"))
```

===

Display a different number of markers at each zoom level using `clusterOptions` on markers or circle markers layers

```{r leaflet_cluster, title = "{{ site.handouts[1] }}"}
leaflet() %>%
  addTiles() %>%
  addMarkers(data = wqp_sites, clusterOptions = markerClusterOptions(),
             popup = ~paste(MntrnLN, ",", OrgnztI))


```

===

Rasters can be slow to load. Be patient but also check out the `maxBytes =` argument if need be. 

```{r leaflet_raster, title = "{{ site.handouts[1] }}"}
leaflet() %>%
  addTiles() %>%
  addRasterImage(nlcd)

```

